Experiences with Planning for Natural Language Generation

### 2009-05-07 trunk/Papers/ci-spark/domains.tex
finished going over 1-3

However, it is desirable to integrate the planner and the generation
system more closely than this. For instance, let's say that the NLG
system wants to generate the instruction sequence ``walk to the centre
of the room; turn right; now press the green button in front of you''.
Experiments \todo{cite} show that this is a pattern that human
instruction givers use frequently: The instruction follower is made to
walk to a certain point in the world at which the instruction giver
can then use a a referring expression (``the green button'') that is
very easy to interpret. \todo{XXX}

However, in order to solve the GIVE task in a more satisfying way, we
expect that planning and generation will have to be interleaved more
closely, making the planning a genuine part of the NLG task. For
instance, the NLG system could generate an instruction sequence
starting with ``turn left; walk forward; press the button; turn right;
walk forward; walk forward; turn right; walk forward; walk forward;
turn left; walk forward''. But such instructions are clumsy and
uninteresting; it would be much better to say ``turn around and walk
through the door''. That is, the system should \emph{summarise} plans
by merging multiple action instances into single instructions before
presenting them to the user.\footnote{This task is similar to the
  construction of ``macro'' actions, but is primarily intended for
  plan presentation purposes. See, e.g., \citep{Botea-etal:05} which
  contains a survey of recent approaches for generating macro actions
  in planning.}  Conversely, it may also be necessary to
\emph{elaborate} on a single planning step by expressing it with
several instructions. Having just entered the top left room, it may be
easier for the user to understand the instruction ``walk to the centre
of the room; turn right; now press the green button in front of you'',
rather than the instruction ``press the green button on the wall to
your right''. Furthermore, in order to plan the referring expression
``the green button in front of you'' at a time when the button is not
yet in front of the user, the NLG system must keep track of the
hypothetical changes in what objects are visible to the user. Thus,
the NLG acts of referring and instructing must be tightly integrated
with the modules for plan generation. This makes the GIVE domain
planning problem an integral part of the NLG system, which makes this
a second planning problem of immediate relevance to NLG that is
completely different from the first one.

% It seems to me that this paragraph misses the point. We're not
% trying to sell GIVE as a cool problem here; we take the perspective
% that GIVE exists and has already been shown to be cool, and we only
% talk about solving it here. - AK
%
% From a planning perspective, GIVE imposes very strict runtime
% requirements on the plan generation task: planning must
% happen in real time and the system must respond to a user in a timely
% fashion.  If the system takes too much time deliberating over an
% instruction to give, rather than actually giving this instruction, the
% user may have walked or turned away, thus making the instruction
% invalid.  Furthermore, plan execution monitoring also plays an
% important role in the GIVE problem.  At a high level, the system needs
% to monitor a user's actions and compare them against the generated
% instruction set to determine if the user has correctly followed
% directions or not. In the case of the latter, new instructions may
% have to be generated. In practice, the situation can be quite
% complicated since the mental state of the user is not known and so the
% system must be inferred from observing the user's actions in real
% time. For example, a user directed to ``turn around and walk through
% the door'' may not necessarily perform these actions to the letter,
% i.e., immediately turning 180 degrees and proceeding directly to the
% door. Instead, the user might take a roundabout route through the
% room, eventually exiting out the door.  Although the user's actions do
% not match the generated instructions exactly, they meet the intended
% goal. The system must be able to identify such ``equivalent'' plans
% and not immediately generate new instructions as soon as the user's
% actions have gone off course. Furthermore, a user can communicate
% certain intentions to the system, both through action and
% inaction. For instance, the system should infer that a user has failed
% to follow instructions if the user exits a room when given a directive
% to ``walk to the centre of the room''. The system should also make a
% similar conclusion if a user simply does nothing when given the
% instruction.

>>>>

However, it is desirable to integrate the planner and the generation
system more closely than this. For instance, let's say that the NLG
system wants to generate the instruction sequence ``walk to the centre
of the room; turn right; now press the green button in front of you''.
Experiments \todo{cite} show that this is a pattern that human
instruction givers use frequently: The instruction follower is made to
walk to a certain point in the world at which the instruction giver
can then use a a referring expression (``the green button'') that is
very easy to interpret. Such an NLG system must integrate discourse
planning and planning in the domain of the world map closely: On the
one hand, the structure of the discourse is determined by the needs of
the NLG system rather than the domain plan; on the other hand, the
discourse planner must be aware of the way in which the instruction
``turn right'' is likely to change the visibility of objects. Even if
the NLG system doesn't implement the generation of this discourse as
planning, it still solves a problem that subsumes the domain planning
problem. For these reasons, we consider the GIVE domain planning
problem as a natural part of a GIVE NLG system.
\todo{Make sure that this paragraph still talks about the things that
  the reviewers found interesting about it.}
